---
title: "Testing pleiotropy vs. separate QTL in multiparental populations"
author: ["<span class=\"presenter\">Frederick J. Boehm</span>, Brian S. Yandell, & Karl W. Broman<ol class=\"affiliations\">University of Wisconsin-Madison, Madison, Wisconsin, USA"]
bibliography: research.bib
output:
  drposter::drposter_poster
---

```{r, echo = FALSE}
options(knitr.chunk.dev = 'svglite')
library(knitr)
```


# {.col-2}


## Introduction

- RNA sequencing and mass spectrometry enable measurement of thousands of phenotypes    
- Multiparental populations enable high-resolution QTL mapping    
- Together, they inform complex trait genetics    
- New analysis tools, such as our test of pleiotropy vs. separate QTL, are needed    

## Background

- @jiang1995multiple developed a pleiotropy vs. separate QTL test for two-parent crosses  
    - Applies to two traits that map to a single genomic region  
    - Null hypothesis: pleiotropy   
    - Alternative: presence of two separate QTL      
    - Perform a two-dimensional QTL scan   
    - Calculate likelihood ratio test statistic  

## Challenges

- Relatedness: *Multivariate polygenic random effects*    
- Eight founder lines: *8 fixed effects*   
- Test statistic calibration: *Parametric bootstrap test*    


## Procedure

- Fit for each marker pair: $$vec(Y) = Xvec(B) + vec(G) + vec(E)$$   
    - $$G \sim MN(0, K, V_g)$$    
    - $$E \sim MN(0, I, V_e)$$    
    - $X$ contains allele probabilities    
    - $B$ contains allele effects    
- Calculate likelihood for each model fit    
- Test statistic: $$- \log_{10} \frac{\max L_0(B, \Sigma, \omega_1)}{\max L_A(B, \Sigma, \omega_1, \omega_2)}$$    
- Parametric bootstrap to get p-value    







## Behavioral genetics

- @logan2013high and @recla2014precise examined 261 Diversity Outbred mice    
    - Identified *Hydin* as the Chr 8 gene affecting "hot plate latency" at 57 cM    
    - Identified Chr 8 QTL for "percent time in light" and "distance traveled in light" at 55 cM    
    - Does *Hydin* affect "percent time in light" and "distance traveled in light"?    




## Allele effects


```{r, echo = FALSE}
include_graphics("~/Box Sync/attie/qtl2pleio-manuscript/Rmd/coef_all3.svg")
```


# {.col-2}


## Profile LODs

```{r, echo = FALSE}
include_graphics("~/Box Sync/attie/qtl2pleio-manuscript/Rmd/all3.svg", auto_pdf = FALSE)
```





## 3 pairwise tests

```{r, echo = FALSE, results='hide'}
options(tidyverse.quiet = TRUE)
library(tidyverse, quietly = TRUE)
```



```{r, echo = FALSE}
tr1 <- c("pct. time in light", "distance traveled", "pct. time in light")
tr2 <- c("hot plate latency", "hot plate latency", "distance traveled")
pv <- c(0.109, 0.108, 0.871)
tibble(Trait1 = tr1, Trait2 = tr2, pvalue = pv) %>%
  kable()
```





## Conclusions

- Evidence for two separate QTL affecting the 3 traits    
    - 1 QTL affects both "distance traveled in light" and "percent time in light"   
    - Second QTL contains *Hydin* and affects "hot plate latency"    

## Future directions

- Examine expression data from @keller2018genetic    
    - 378 Diversity Outbred mice    
    - Expression QTL hotspot dissection    
    - Statistical power studies    
    - Comparisons with mediation analyses    


## Contact information

Fred Boehm  
EMAIL: frederick.boehm@gmail.com  
WEBSITE: https://fboehm.us/  
ORCID: 0000-0002-1644-5931  
`qtl2pleio` R package: https://github.com/fboehm/qtl2pleio    


## References


